Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-03-29 paper/1SUB/intro.tex
saved old version

This paper is the first model of relaxed memory with a compositional
semantics for sequential composition.  It shows how pomsets with
preconditions~\cite{DBLP:journals/pacmpl/JagadeesanJR20} can be combined
with predicate transformers~\cite{DBLP:journals/cacm/Dijkstra75}.
\begin{itemize}
\item \S\ref{sec:model} presents the basic model, with few features
  required of the logic of preconditions, but a resulting lack of fidelity
  to exiting models,
\item \S\ref{sec:q} adds a model of \emph{quiescence} to the logic,
  required to model coherence (accessing $x$ has a precondition that $x$ is quiescent)
  and synchronization (a releasing write requires all locations to be quiescent),
\item \S\ref{sec:arm} adds the features required for efficient compilation
  to modern architectures: downgrading some synchronized accesses to relaxed,
  and removing read-read dependencies, and
\item \S\ref{sec:complications} show how to address common litmus tests.
\end{itemize}
% Acknowledgements go here, once we're not double-blinded.
The definitions in this paper have been formalized in Agda.

@@

As far as we are aware, there are no previous attempts to provide a
compositional semantics of sequential composition in a relaxed memory model.
For a discussion of related work for relaxed memory models in general, see
\cite{DBLP:journals/pacmpl/JagadeesanJR20}.

>>>>

In this paper, we show how pomsets with
preconditions~\cite{DBLP:journals/pacmpl/JagadeesanJR20} can be combined with
predicate transformers~\cite{DBLP:journals/cacm/Dijkstra75} to create a
compositional semantics for sequential composition.
\begin{itemize}
\item \S\ref{sec:model} presents the basic model, with few features
  required of the logic of preconditions, but a resulting lack of fidelity
  to exiting models,
\item \S\ref{sec:q} adds a model of \emph{quiescence} to the logic,
  required to model coherence (accessing $x$ has a precondition that $x$ is quiescent)
  and synchronization (a releasing write requires all locations to be quiescent),
\item \S\ref{sec:arm} adds the features required for efficient compilation
  to modern architectures: downgrading some synchronized accesses to relaxed,
  and removing read-read dependencies, and
\item \S\ref{sec:complications} show how to address common litmus tests.
\end{itemize}
% Acknowledgements go here, once we're not double-blinded.
The definitions in this paper have been formalized in Agda.

@@

\section{Related Work}

\begin{itemize}
\item Models with strong dependencies that disallow compiler optimizations
  and efficient implementation on \armeight{}:
  \cite{Lamport:1979:MMC:1311099.1311750,
    DBLP:conf/pldi/LahavVKHD17,
    Dolan:2018:BDR:3192366.3192421,
    DBLP:conf/pldi/LahavVKHD17,
    DBLP:conf/lics/JeffreyR16,
    Boehm:2014:OGA:2618128.2618134,
    DBLP:journals/corr/abs-1804-04214}.
\item Models with weak dependencies that allow some forms of thin-air
  reasoning:
  \cite{Manson:2005:JMM:1047659.1040336,
    DBLP:conf/esop/JagadeesanPR10,
    DBLP:conf/popl/KangHLVD17,
    DBLP:journals/pacmpl/ChakrabortyV19}
\item Models that do not calculate dependencies:
  \cite{Batty:2011:MCC:1926385.1926394,
    DBLP:journals/pacmpl/WattRP19,
    DBLP:conf/pldi/WattPPBDFPG20,
    goMM}.
\end{itemize}
Like us, \citet{DBLP:journals/corr/abs-1804-04214} use pomsets to create a
compositional account of sequential composition.  However, their model
requires a fence after every read on \armeight{}, and uses syntactic
dependencies, thus invalidating many compiler optimizations.
